---
title: "Untitled"
output: html_document
date: '2022-07-15'
---

## VII. APA Tables & Results Exporting

```{r}
# Model summary table
table1=APAtable_Bayesian_LogReg(Thesis_Model)

docx_file <- here::here("Results Output",(fileext = "new table.docx"))

save_as_docx(table1, path = docx_file)
```


## EXTRACTING \_\_\_ FROM THE\_\_\_

### ...draws from the posterior

-   These commands draw samples of *parameter values* from the posterior.

-   Kurz uses `brms::posterior_samples` in his translation books but this is DEPRICATED. Best bet is to use `tidy_draws`; see [here](SEE%20https://mjskay.github.io/tidybayes/articles/tidy-brms.html#introduction) for info. The {tidybayes} package follows the tidyverse format.

```{r}
insight::get_parameters(b4.3) |> head() # comparable to posterior_samples, except it doesn't give you samples for lp__

tidybayes::tidy_draws(b4.3) |> head() # USE THIS ONE. {tidybayes} is useful for extracting draws in a form and manner consistent w/ the tidyverse.
```

### ...draws from the PRIOR Predictive Distribution

-   When performing prior predictive checks (below), the following command can be used to extract values from the prior to show what the model expects, *before* seeing the observed data.

```{r}
brms::prior_draws()
```

### ...fitted (i.e., predicted) values from the POSTERIOR Predictive Distribution

-   This is a method that uses the posterior to generate data that would be expected by the model. The results of these commands return *the data your model would expect to see,* given the parameters implied in the posterior.

> With `fitted()`, we compute the expected values for each state.

```{r}
fitted(b5.4) |> as_tibble() |> head()
```

Kurz extracts the fitted values (values expected by the model) and then binds this back in with the full data

```{r}
fitted(b5.4) |> data.frame() |> bind_cols(d) |> head()
```

### Residuals
```{r}
residuals(b5.2) |> as_tibble()
```



## BAYESIAN PLOTS AND FIGURES

### Bayes factors, ROPE's, posteriors, etc.

<https://easystats.github.io/see/articles/bayestestR.html>

-   The {see} and {bayestestR} packages making plotting Bayesian models very easy. General format for almost everything is the same...

-   To plot just the thing, combine the desired {bayestestR} command with `plot`; for example:

```{r}

# Significance testing with ROPE
ropetest=rope(Thesis_Model,range = c(-0.181,0.181), ci=0.89, ci_method = "HDI")

result <- equivalence_test(Thesis_Model)
plot(result)


plot(bayestestR::rope(b5.3))
```

To customize the look and appearance of the plot, use the same format as above, and add ggplot themes or themes and colors from other packages:

```{r}

#HDI visuals
CIs <- hdi(Thesis_Model, ci = c(0.89))
plot(CIs) + scale_fill_got_d(option = "white_walkers")

plot(bayestestR::hdi(b5.3)) + 
  theme_classic()+
  scale_fill_metro(palette = "ice")


#point estimates
result <- point_estimate(Thesis_Model, centrality = c("map"))
result
plot(result, panel = FALSE)[[1]]


# Probability of Direction
plot(pds)+
  scale_fill_manual(values=c("#FFC107", "#E91E63"))+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "italic"))

plot(p_direction(model, priors=FALSE))

plot(bayestestR::pd(b5.3)) + 
  theme_classic()+
  scale_fill_metro(palette = "ice")

```

### Coefficient estimates

See this page for help: <https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html>

```{r}
# Simple quick-and-dirty option
mcmc_plot(b5.2, pars = "^b_")


# fancy version

library(bayesplot)

color_scheme_set("red")

post <- tidybayes::tidy_draws(b5.9)

post %>% 
  select(starts_with("b_")) %>% 
  
  bayesplot::mcmc_intervals(prob = .5, point_est = "median") + # the key plotting feature here
  
  labs(title = "My fancy coefficient plot") +
  theme_bw() +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank())
```
